Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-01-09 eswc-2012/phdsymp.tex
More disposition
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@49 43ef1190-90fd-4dd7-a769-d0121f02b80e
Elevator pitches
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@50 43ef1190-90fd-4dd7-a769-d0121f02b80e

Query federation with SPARQL has attracted much attention
from industry and academia alike, and four implementations of basic
query federation were submitted to the SPARQL 1.1 Working Group as
input for the forthcoming work. The basic query federation feature was
supported by a large number of group members, and the latest working
draft of the proposed standard was published on June 1st 2010 and is
expected to enter a Last Call review period shortly. % TODO: Check
                                % status

@@

Since the main objective of the proposed work is to create systems
that have sufficient performance for practical applications, it is of
paramount importance to have methodology that can falsify a
conjecture about an implementation's performance.

@@

@@

@@

>>>>

Query federation with SPARQL has attracted much attention
from industry and academia alike, and four implementations of basic
query federation were submitted to the SPARQL 1.1 Working Group as
input for the forthcoming
work\footnote{http://www.w3.org/2009/sparql/wiki/Feature:BasicFederatedQuery}. 
The basic query federation feature was
supported by a large number of group members, and the Last Call working
draft of the proposed standard was published on 17 November 2011.

@@

Since the main objective of the proposed work is to create systems
that have sufficient performance for practical applications, it is of
paramount importance to have methodology that can falsify a
hypothesis about an implementation's performance.

@@

\subsection{In SPARQL Federation}

This part of the thesis addresses problems of query optimization in a
federated regime.

\subsection{In Benchmarking}

The contemporary problem in benchmarking is to distinguish
implementations when the measured difference is little or susceptible
to random noise.

@@

\subsection{In SPARQL Federation}

\subsubsection{Technology}

I take the technology state of the art to be represented by the
current basic SPARQL 1.1 Federated Query Working
Draft\footnote{http://www.w3.org/TR/2011/WD-sparql11-federated-query-20111117/}. For
an academic treatment of the current specification, see
\cite{springerlink:10.1007/978-3-642-21064-8_1} and references therein.

\subsubsection{Science}

In \cite{springerlink:10.1007/978-3-642-21064-8_1}, the authors show
an optimization strategy based on execution order of different
algebraic elements. % TODO: wording here

In \cite{5337556}, it is shown how histograms can be computed to
assist the SemWIQ optimizer, written by the same authors.
% RDFStats

\subsection{In Benchmarking}

Numerous benchmarks have been developed for SPARQL, but
\cite{Duan:2011:AOC:1989323.1989340} showed that currently, most
benchmarks poorly represent the typical data and queries that is used
on the Semantic Web.

Most recently, \cite{mxro:Morsey2011DBpedia} addressed some of these
problems by using real data and real queries from DBpedia.

The problems addressed by these studies are almost orthogonal to the
problems considered by my proposed project.

@@

\subsection{In SPARQL Federation}

There are many possible approaches for this part of the thesis. At
this stage, it seems that use of selectivity estimation may be the
most feasible direction.

\subsection{In Benchmarking}

This part of the thesis seeks to use advances from statistical
experimental planning to improve accuracy and dependability of benchmarking.
